---
title: "Depth interval estimate vs Point-depth estimates"
author: "Stephen Roecker"
date: "7/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}

library(aqp)
library(sp)
library(GSIF)

```


This example dataset is from the GSIF R package help file for the mpspline().

```{r, message=FALSE}

## sample profile from Nigeria:
prof1 <- data.frame(id = "ISRIC:NG0017", 
                    top = c(0, 18, 36, 65, 87, 127), 
                    bot = c(18, 36, 65, 87, 127, 181),
                    ORCDRC = c(18.4, 4.4, 3.6, 3.6, 3.2, 1.2),
                    munsell = c("7.5YR3/2", "7.5YR4/4", "2.5YR5/6", "5YR5/8", "5YR5/4", "10YR7/3")
                    )
loc <- data.frame(lon = 3.90, lat = 7.50, id = "ISRIC:NG0017", FAO1988 = "LXp")
prof1 <- merge(prof1, loc, by = "id", all.x = TRUE)

depths(prof1) <- id ~ top + bot
site(prof1) <- ~ lon + lat + FAO1988 
coordinates(prof1) <- ~ lon + lat
proj4string(prof1) <- CRS("+proj=longlat +datum=WGS84")

# fit a spline:
prof1_mps <- GSIF::mpspline(prof1, var.name="ORCDRC")
prof1_mps$var.std            # depth interval (e.g. block mean)
head(prof1_mps$var.1cm[, 1]) # depth slices   (e.g. point estimates)


# aqp depth slice approach
prof1_slab  <- aqp::slab(prof1, id ~ ORCDRC, slab.structure = c(0, 5, 15, 30, 60, 100, 200), slab.fun = mean)
prof1_slice <- aqp::slice(prof1, 0:200 ~ ORCDRC)

```

The below example confirms that the var.std estimates are the mean of the var.1cm for the 0-5 cm depth interval.

```{r}

# 0-5 cm 
# spline approach - depth interval
prof1_mps$var.std[1]
mean(prof1_mps$var.1cm[1:5])

```

The below example confirms that the slab() depth interval estimate is the same as the mean of the slice() point depth estimates for the 0-5 cm depth interval.


```{r}

# slice approach - depth interval
prof1_slab$value[1]
mean(horizons(prof1_slice)$ORCDRC[1:5])

```


True (?) point depth estimates for the spline and slice approaches.

```{r}

idx <- round(c(mean(0:5), mean(5:15), mean(15:30), mean(30:60), mean(60:100), mean(100:200)))
idx

# spline approach - depth slice
prof1_mps$var.1cm[idx]


# slice approach - depth slice
horizons(prof1_slice)$ORCDRC[idx]

```

