---
title: "Depth interval estimate vs Point-depth estimates"
author: "Stephen Roecker"
date: "7/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}

library(aqp)
library(soilDB)
library(mpspline2)
library(collapse)

```


Example of mpspline R function.

```{r, message=FALSE}

## sample profile
dat <- fetchKSSL("Miami")
sub <- horizons(dat[28])[c("pedon_key", "hzn_top", "hzn_bot", "clay")]
list(sub$clay, sum(sub$clay))

spc <- sub
depths(spc) <- pedon_key ~ hzn_top + hzn_bot


# fit a spline:
dep <- c(0, 25, 36, 61, 86, 112, 152)

## if lam = 0 and spline depths = horizon depth, then the predicted spline values should be close/equal to the observed input values for each horizon
{mpspline(sub, var_name = "clay", d = dep, lam = 0)$`38259`$est_dcm ->.;
  list(values = round(., 1), sum = sum(.) |> round(1))
}

## if the low and high values are set, then the spline predictions are constrainted
{mpspline(sub, var_name = "clay", d = dep, lam = 0, vlow = min(sub$clay), vhigh = max(sub$clay))$`38259`$est_dcm ->.;
  list(values = round(., 1), sum = sum(.) |> round(1))
}

## if the default is lamda = 0.1, then the overall spline prediction sum equals the orignal values sum
{mpspline(sub, var_name = "clay", d = dep)$`38259`$est_dcm ->.;
  list(values = round(., 1), sum = sum(.) |> round(1))
}


# aqp depth slice approach
prof1_slab  <- aqp::slab(spc, pedon_key ~ clay, slab.structure = c(0, 5, 15, 30, 60, 100, 200), slab.fun = mean)
prof1_dice <- aqp::dice(spc, 0:200 ~ clay)

prof_seg <- segment(spc, intervals = c(0, 5, 15, 30, 60, 100, 200)) |>
  transform(thk = hzn_bot - hzn_top) |>
  horizons()
{collapse::fmean(prof_seg$clay, g =  prof_seg$segment_id, w = prof_seg$thk) |> round(1) ->.;
  list(round(., 1), sum(.) |> round(1))
  }

```

The below example confirms that the var.std estimates are the mean of the var.1cm for the 0-5 cm depth interval.

```{r}

# 0-5 cm 
# spline approach - depth interval
test <- mpspline(sub, var_name = "clay")
test$`38259`$est_dcm
mean(test$`38259`$est_1cm[1:5])

```

The below example confirms that the slab() depth interval estimate is the same as the mean of the slice() point depth estimates for the 0-5 cm depth interval.


```{r}

# slice approach - depth interval
prof1_slab$value[1]
mean(horizons(prof1_dice)$clay[1:5])

```


True (?) point depth estimates for the spline and slice approaches.

```{r}

idx <- round(c(mean(0:5), mean(5:15), mean(15:30), mean(30:60), mean(60:100), mean(100:200)))
idx

# spline approach - depth slice
prof1_mps$var.1cm[idx]


# slice approach - depth slice
horizons(prof1_slice)$ORCDRC[idx]

```


# plot example of methods

```{r}

library(soilDB)
library(aqp)
library(mpspline2)
library(ggplot2)


vars <- c("pedon_key", "hzn_top", "hzn_bot", "clay")
int <- c(0, 5, 15, 30, 60, 100, 200)
mid <- round(rowMeans(cbind(int, c(int[-1], NA))))
int2 <- sort(c(mid, mid + 1))


f    <- fetchKSSL("Miami")
idx  <- site(f)$pedon_id %in% "S2012OH021001"
test <- f[idx, ]
test2 <- segment(test, intervals = 0:150)
test_df <- cbind(group = "original data", horizons(test2)[vars[-3]])


# point estimate
f_p <- segment(test, intervals = int2)
idx <- which(horizons(f_p)$hzn_top %in% mid)
f_p <- f_p[, idx]
pd_df <- cbind(group = "point sample", horizons(f_p)[vars[-3]])


# spline estimate
test_s <- mpspline(horizons(test)[vars], var_name = "clay", d = int, vlow = min(test$clay), vhigh = max(test$clay))
mps_df <- data.frame(group = "spline estimate",
                     pedon_key = test_s[[1]]$pedon_key, 
                     hzn_top   = 1:length(test_s[[1]]$est_1cm), 
                     clay      = test_s[[1]]$est_1cm
                     )
# mps_df <- mps_df[mps_df$hzn_top %in% mid, vars[-3]]


# combine
test_c <- rbind(test_df, pd_df, mps_df)

col = c(blue = "#0000FF", black = "#17202A", orange = "#F39C12")

# plot
ggplot(test_df, aes(col = group)) +
  # geom_line(data = test_df, aes(y = clay, x = hzn_top, col = group), lwd = 2) +
  geom_line(data = mps_df, aes(y = clay, x = hzn_top, group = "spline"), lwd = 2, linetype = "longdash") +
  # geom_point(data = pd_df, aes(y = clay, x = hzn_top), size = 3.5) +
  scale_x_continuous(breaks = int) +
  # scale_color_manual(values = c("blue", "black", "orange")) +
  scale_color_manual(values = unname(col)[3]) +
  coord_flip(xlim = c(150, 0)) + 
  #xlim(150, 0) +
  xlab("depth interval (cm)") + ylab("clay (%)") +
  ggtitle("Comparion of methods") + 
  theme_bw()
ggsave(dp_gg, filename = "test.png", device = "png", width = 6, height = 5)

```
